Notes on Gravitational Entropy

### 2021-05-16 alpha/GravitationalEntropy/entropy.tex
Update note on gravitational entropy

	Although we've expressed everything in terms of the bulk parameters and the worldsheet deformation parameter $\lambda$, we'd like to emphasize that the above results comes from purely field theoretical calculations, and they are mapped to the bulk using the holographic dictionary. Our goal is to reproduce them as gravitational entropy, from bulk calculations. 
\subsection{Case study: Ramond vacuum $T_u = T_v = 0$}
	For simplicity, we first look at the \textsl{Ramond vacuum}, i.e.~the TsT background with $T_u = T_v = 0$. In this case the string frame metric and dilaton is given by:
	\begin{equation}
		\dd{s}^2
		= \frac{\dd{z}^2}{z^2}
			+ \frac{\dd{u}\dd{v}}{z^2 + 2\lambda},
	\quad
		e^{2\Phi}
		= \frac{k}{p}
			\frac{z^2}{z^2 + 2\lambda}\,
			e^{-2\phi_0},
		\label{eq:staticMetricPoincare}
	\end{equation}
	To avoid confusion with the dilaton, we denote the angular direction with the curvy $\varphi = \frac{u + v}{2}$. 
	This is derived from (3.13) of \cite{Apolo:2019zai}, with $r = \frac{1}{z^2}$; we've set the AdS radius $\ell_{\mrm{AdS}} = 1$ for now and neglected the $B$ field. For more simplifications, we further set:
	\begin{equation}
		p = k = 1
	\end{equation}
\subsubsection{RT surface}
	The Einstein frame metric $G_{\mu\nu}$ is related to the string frame metric $g_{\mu\nu}$ by:
	\begin{equation}
		G_{\mu\nu} = e^{-4\Phi} g_{\mu\nu}
	\end{equation}
	Now consider a spacelike geodesic in Einstein frame within the time slice $t = 0$; we can nicely parametrize it  with its peak $z_0$ at $\varphi = 0$. The $\pdd{\varphi}$ symmetry implies that along the geodesic $\gamma$,
	\begin{equation}
		\pqty{\dv{z}{\varphi}}_{\!\gamma}^{\!2}
		= \frac{z_0^2}{z^2}
				\frac{z_0^2}{z_0^2 + 2\lambda}
			- \frac{z^2}{z^2 + 2\lambda}
	\end{equation}

	The coordinate length $L$ at the boundary is given by:
	\begin{equation}
	\begin{aligned}
	  L = L(z_0;\mu)
	  &= \int \dd{\varphi}
	  = 2\int_0^{z_0}
	      \abs{\dv{\varphi}{z}}
	      \dd{z} \\
	  &= 2\int_0^{z_0}
	      \frac{1}{\sqrt{
	          \frac{z_0^2}{z^2}
	          \frac{z_0^2}{z_0^2 + 2\lambda}
	          - \frac{z^2}{z^2 + 2\lambda}
	      }}
	      \dd{z}
	\end{aligned}
	\label{eq:length_function}
	\end{equation}

\pagebreak\noindent
	Note that:
	\begin{itemize}
	\item $\mu\to 0,\, L(z_0;\mu) \to 2z_0$, we recover the AdS result: in this case the geodesic $\gamma$ is a semi-circle with radius $z_0$.

	\item $z_0\to 0,\,
		L(z_0;\mu) \to \sqrt{\frac{\pi^2}{2} \lambda}
	$, we see that $L$ is bounded from below by:
	\begin{equation}
		L_{\min} (\mu)
		= \sqrt{\frac{\pi^2}{2} \lambda}
	\end{equation}
	This feature is very similar to the lower bound $L_0$ of entanglement length, but they do \textit{not} agree; in fact,
	\begin{equation}
		L_{\min} / L_0 = \pi / 4
	\end{equation}
	\end{itemize}

	The entanglement entropy, as proposed by RT, is given by:
	\begin{equation}
	\begin{aligned}
	  S_\mrm{RT}
	  &= \frac{1}{4G} \int_\gamma
	    \frac{z^2 + 2\lambda}{z^2}
	    \sqrt{
	        \frac{\dd{z}^2}{z^2}
	        + \frac{\dd{\varphi}^2}{z^2 + 2\lambda}
	    } \\
	  &= \frac{1}{2G} \int_0^{z_0} \dd{z}
	    \frac{z^2 + 2\lambda}{z^2}
	      \sqrt{
	          \frac{1}{z^2}
	          + \frac{1}{z^2 + 2\lambda}
	          \Big/
	          \pqty{\dv{z}{\varphi}}^{\!\!2}
	      } \\
	  &= \frac{1}{2G} \int_0^{z_0} \dd{z}
	    \frac{z^2 + 2\lambda}{z^2}
	      \sqrt{
	          \frac{1}{z^2}
	          + \frac{1}{z^2 + 2\lambda}
	          \Big/
	          \pqty{
					\frac{z_0^2}{z^2}
					\frac{z_0^2}{z_0^2 + 2\lambda}
					- \frac{z^2}{z^2 + 2\lambda}
	          }
	      }
	\end{aligned}
	\end{equation}
	We don't have an explicit analytic result for the above integral, but numerical calculations suggest that it does not agree with \eqref{eq:tst_ft_ee}. 
	As is inspired by the cutoff picture, we would like to tweak it such that it matches with \eqref{eq:tst_ft_ee}.

	We have the following observations: 
	\begin{itemize}
	\item If we denote $L_{\mrm{CFT}} = 2z_0$, and instead of using \eqref{eq:length_function}, we use:
	\begin{equation}
		L' = 2\sqrt{z_0^2 + 2\lambda},
	\quad\text{i.e.}\quad
		\pqty{\frac{L'}{2}}^{\!\!2}
		= \pqty{\frac{L_\mrm{CFT}}{2}}^{\!\!2} + 2\lambda,
	\end{equation}
	Which mimic the relation \eqref{eq:length_relations}, then the $C(L)$ from the field theory calculations can be re-written as:
	\begin{equation}
		C(L')
		= \frac{L'}{L_{\mrm{CFT}}}
		= \frac{1}{\sqrt{1 - (\frac{L_0}{L'})^2}},
	\quad L_0 = \sqrt{8\lambda}
	\end{equation}
	Where the relation $C(L') = \frac{L'}{L_{\mrm{CFT}}}$ is completely analogous to the double-trace situation; in that case we have $L_\mrm{CFT} = L_\infty$.

@@

	\item Therefore, we have:
	\begin{equation}
		L' \pdv{S_\mrm{RT}}{L'}
		= L_\mrm{CFT} \pdv{S_\mrm{RT}}{L_\mrm{CFT}}
			\pdv{L_\mrm{CFT}}{L'}
			\frac{L'}{L_\mrm{CFT}}
		\sim \frac{1}{2G}
			\pqty{\frac{L'}{L_\mrm{CFT}}}^{\!\!2}
		= C(L')^2
	\end{equation}
	We find that \textit{numerically}, this does hold very well. Unfortunately this is not what we expected. How should we interpret this?
	\end{itemize}

	\begin{figure}[!ht]
	\centering
	\includegraphics[width=.7\linewidth]{img/cFunctionGame.pdf}
%	\hspace{2em}
%	\vspace{-2ex}
	\caption[Numerical construction of the $C(L)$ function]{%
		Numerical construction of the $C(L)$ function from $S_\mrm{RT}$ by considering $
			\sqrt{
				\frac{1}{2G}
				L' \pdv{S}{L'}
			}
		$. What's happening? I have no idea.
	}
	\end{figure}

>>>>

	Although we've expressed everything in terms of the bulk parameters and the worldsheet deformation parameter $\lambda$, we'd like to emphasize that the above results comes from purely field theoretical calculations, and they are mapped to the bulk using the holographic dictionary. Our goal is to reproduce them as gravitational entropy, from bulk calculations. 
\subsection{Case study: Ramond vacuum $T_u = T_v = 0$}
	For simplicity, we first look at the \textsl{Ramond vacuum}, i.e.~the TsT background with $T_u = T_v = 0$. In this case the string frame metric and dilaton is given by:
	\begin{equation}
		\dd{s}^2
		= \frac{\dd{z}^2}{z^2}
			+ \frac{\dd{u}\dd{v}}{z^2 + 2\lambda},
	\quad
		e^{2\Phi}
		= \frac{k}{p}
			\frac{z^2}{z^2 + 2\lambda}\,
			e^{-2\phi_0},
		\label{eq:staticMetricPoincare}
	\end{equation}
	To avoid confusion with the dilaton, we denote the angular direction with the curvy $\varphi = \frac{u + v}{2}$. 
	This is derived from (3.13) of \cite{Apolo:2019zai}, with $r = \frac{1}{z^2}$; we've set the AdS radius $\ell_{\mrm{AdS}} = 1$ for now and neglected the $B$ field. For more simplifications, we further set:
	\begin{equation}
		p = k = 1
	\end{equation}
\subsubsection{The usual RT surface}
	The Einstein frame metric $G_{\mu\nu}$ is related to the string frame metric $g_{\mu\nu}$ by:
	\begin{equation}
		G_{\mu\nu} = e^{-4\Phi} g_{\mu\nu}
	\end{equation}
	Now consider a spacelike geodesic in Einstein frame within the time slice $t = 0$; we can nicely parametrize it  with its peak $z_0$ at $\varphi = 0$. The $\pdd{\varphi}$ symmetry implies that along the geodesic $\gamma$,
	\begin{equation}
		\pqty{\dv{z}{\varphi}}_{\!\gamma}^{\!2}
		= \frac{z_0^2}{z^2}
				\frac{z_0^2}{z_0^2 + 2\lambda}
			- \frac{z^2}{z^2 + 2\lambda}
	\end{equation}

	The coordinate length $L_\mrm{RT}$ at the boundary is given by:
	\begin{equation}
	\begin{aligned}
	  L_\mrm{RT} = L_\mrm{RT}(z_0;\lambda)
	  &= \int \dd{\varphi}
	  = 2\int_0^{z_0}
	      \abs{\dv{\varphi}{z}}
	      \dd{z} \\
	  &= 2\int_0^{z_0}
	      \frac{1}{\sqrt{
	          \frac{z_0^2}{z^2}
	          \frac{z_0^2}{z_0^2 + 2\lambda}
	          - \frac{z^2}{z^2 + 2\lambda}
	      }}
	      \dd{z}
	\end{aligned}
	\label{eq:length_function}
	\end{equation}
%	
%\pagebreak\noindent
	Note that:
	\begin{itemize}
	\item $\lambda\to 0$, \textsl{or} $z_0\to \infty$, we have $L_\mrm{RT}(z_0;\lambda) \to 2z_0$, i.e.~we recover the AdS result: in this case the geodesic $\gamma$ is a semi-circle with radius $z_0$.

	\item $z_0\to 0,\,
		L_\mrm{RT}(z_0;\lambda) \to \sqrt{\frac{\pi^2}{2} \lambda}
	$, we see that $L_\mrm{RT}$ seems to be bounded from below by:
	\begin{equation}
		L_{\mrm{RT},0} (\lambda)
		= \sqrt{\frac{\pi^2}{2} \lambda}
	\end{equation}
	This feature is very similar to the lower bound $L_0$ of entanglement length, but they do \textit{not} agree; in fact,
	\begin{equation}
		L_{\mrm{RT},0} / L_0 = \pi / 4
	\end{equation}
	\end{itemize}

	The entanglement entropy, as proposed by RT, is given by:
	\begin{equation}
	\begin{aligned}
	  S_\mrm{RT}
	  &= \frac{1}{4G} \int_\gamma
	    \frac{z^2 + 2\lambda}{z^2}
	    \sqrt{
	        \frac{\dd{z}^2}{z^2}
	        + \frac{\dd{\varphi}^2}{z^2 + 2\lambda}
	    } \\
	  &= \frac{1}{2G} \int_0^{z_0} \dd{z}
	    \frac{z^2 + 2\lambda}{z^2}
	      \sqrt{
	          \frac{1}{z^2}
	          + \frac{1}{z^2 + 2\lambda}
	          \Big/
	          \pqty{\dv{z}{\varphi}}^{\!\!2}
	      } \\
	  &= \frac{1}{2G} \int_0^{z_0} \dd{z}
	    \frac{z^2 + 2\lambda}{z^2}
	      \sqrt{
	          \frac{1}{z^2}
	          + \frac{1}{z^2 + 2\lambda}
	          \Big/
	          \pqty{
					\frac{z_0^2}{z^2}
					\frac{z_0^2}{z_0^2 + 2\lambda}
					- \frac{z^2}{z^2 + 2\lambda}
	          }
	      }
	\end{aligned}
	\label{eq:RT_integral}
	\end{equation}
	We don't have an explicit analytic result for the above integral\footnote{
		Note that there are some analytic expressions given by \cite{Asrat:2019end}, but we haven't confirmed their results. 
		We can map \cite{Asrat:2019end} to our convention by matching the aysmptotic behavior of $L_\mrm{RT}$; namely, we have $z_0 \to \infty$, $L_\mrm{RT}(z_0;\lambda)\to 2z_0$, while $z_0 \to 0$, $L_\mrm{RT}(z_0;\lambda)\to \sqrt{\frac{\pi^2}{2} \lambda}$. It seems that for (3.17) and (4.3) in \cite{Asrat:2019end}, we have $
			\alpha \mapsto \frac{\lambda/2}{z^2}
		$. With this the length $L_\mrm{RT}$ given by \cite{Asrat:2019end} as a function of $z$ is \textit{not} monotonic, which sounds suspicious and does not agree with numerical results. \label{foot:asrat_matching}
	},
	but numerical calculations suggest that it does not agree with \eqref{eq:tst_ft_ee}. \sidenote{However, its qualitative behavior is quite close to \eqref{eq:tst_ft_ee}, and it fits \eqref{eq:tst_ft_ee} nicely as $L\to \infty$. }

\pagebreak[3]

	As is inspired by the cutoff picture, we would like to tweak the holographic entropy such that it matches with \eqref{eq:tst_ft_ee}. 
	We start with the following observations: 
	\begin{itemize}
	\item If we denote $L_{\mrm{CFT}} = 2z_0$, and instead of using \eqref{eq:length_function}, we use:
	\begin{equation}
		L' = 2\sqrt{z_0^2 + 2\lambda},
	\quad\text{i.e.}\quad
		\pqty{\frac{L'}{2}}^{\!\!2}
		= \pqty{\frac{L_\mrm{CFT}}{2}}^{\!\!2} + 2\lambda,
	\end{equation}
	Which mimic the relation \eqref{eq:length_relations}, then the $C(L)$ from the field theory calculations can be re-written as:
	\begin{equation}
		C(L')
		= \frac{1}{\sqrt{1 - (\frac{L_0}{L'})^2}}
		= \frac{L'}{L_{\mrm{CFT}}},
	\quad L_0 = \sqrt{8\lambda}
	\end{equation}
	Where the relation $C(L') = \frac{L'}{L_{\mrm{CFT}}}$ is completely analogous to the double-trace situation; in that case we have $L_\mrm{CFT} = L_\infty$.

	\pagebreak[3]

	\item Maybe one can ``engineer'' a suitable curve $\gamma'$ which has peak $z_0$ and boundary length $
		L' = 2\sqrt{z_0^2 + 2\lambda}
	$; we can require that it reproduces the desired entanglement entropy \eqref{eq:tst_ft_ee}:
	\begin{equation}
		\frac{1}{2G} \mop{arccosh}
			\frac{
				\sqrt{z_0^2 + 2\lambda}
			}{\sqrt{2\lambda}}
	  = \frac{1}{2G} \int_0^{z_0} \dd{z}
	    \frac{z^2 + 2\lambda}{z^2}
	      \sqrt{
	          \frac{1}{z^2}
	          + \frac{1}{z^2 + 2\lambda}
	          \pqty{\dv{\varphi}{z}}^{\!\!2}
	      } 
	\end{equation}
	This is a integral equation for $\dv{\varphi}{z} \equiv F$. The above equation can be greatly simplified by:
	\begin{equation}
		z = \sqrt{2\lambda} \sinh y,
	\quad
		F(y;y_0,\lambda)
		= \abs{\dv{\varphi}{z}},
	\end{equation}
	\begin{empheq}{equation}
		y_0
		= \int_0^{y_0} \dd{y}
			\coth^2 y\,\sqrt{
				\coth^2 y
				+ F^2(y;y_0,\lambda)
			}
	\end{empheq}

	$z_0$ or $y_0$ labels the peak of $\gamma$, thus we have:
	\begin{empheq}{equation}
		F(y\to y_0;y_0,\lambda) \to \infty
	\end{empheq}
	Also the boundary length should be:
	\begin{gather}
	  L' = 2\sqrt{z_0^2 + 2\lambda}
	  = 2\int_0^{z_0}
	      \abs{\dv{\varphi}{z}}
	      \dd{z},
	\end{gather}
	\begin{empheq}{equation}
		\cosh y_0
		= \int_0^{y_0}
			F(y;y_0,\lambda) \cosh y \dd{y}
	\end{empheq}

	In principle one might be able to find some candidate $F = F(y;y_0,\lambda)$ from the boxed equations. This corresponds to candidate curve $\gamma'$ that reproduces the desired entropy.

	\item There is a $B$-field in the TsT background; however, as is argued in \texttt{areaTTbar.pdf}, it should not contribute to the on-shell action. However, one might be able to promote the original $\gamma_\mrm{RT}$ to a tube surrounding $\gamma_\mrm{RT}$, and treat it as a string worldsheet, where the $B$-field couples nicely in the worldsheet action.

	There are two ways going forward: one is to compute the string EoMs which is related to the \textit{string frame} geodesic; this is done in \texttt{notesEETTbar.pdf} and to be discussed in the next section.

	\sidenote{Another way is to come up with a procedure that reduces the string action to a particle action close to $S_{\mrm{RT}}$, but with $B$-field corrections. One can also construct a new connection with $H = \dd{B}$ and solve the geodesic with this new connection\footnote{
		Thank Yuan Zhong for this great idea. Also he helped a lot with simplying the integral equations. 
	}. }

@@

	However, if we use the analytic result of \cite{Asrat:2019end} (see footnote \ref{foot:asrat_matching}), then this is \textit{not} correct! But one should check that result first to see if it's valid first.
	\sidenote{Also,} if we take the $z_0$ derivative before integrating in \eqref{eq:RT_integral}, we can show that indeed we have, approximately, $
		z_0\,\pdv{z_0}\,S_\mrm{RT}
		\approx \frac{1}{2G}
	$, but the result is \sidenote{\textit{not} exactly constant and \textit{not} $\lambda$ independent;} see \texttt{zDzNonConstant.nb}.

\pagebreak[3]

	If we proceed anyways with $
		z_0\,\pdv{z_0}\,S_\mrm{RT}
		\approx \frac{1}{2G}
	$, then we have:
	\begin{equation}
		L' \pdv{S_\mrm{RT}}{L'}
		= L_\mrm{CFT} \pdv{S_\mrm{RT}}{L_\mrm{CFT}}
			\pdv{L_\mrm{CFT}}{L'}
			\frac{L'}{L_\mrm{CFT}}
		\approx \frac{1}{2G}
			\pqty{\frac{L'}{L_\mrm{CFT}}}^{\!\!2}
		= C(L')^2
	\end{equation}
	We find that \textit{numerically}, this does hold very well. Unfortunately this is not what we expected. How should we interpret this?
	\end{itemize}

	\begin{figure}[!ht]
	\centering
	\includegraphics[width=.7\linewidth]{img/cFunctionGame.pdf}
%	\hspace{2em}
%	\vspace{-2ex}
	\caption[Numerical construction of the $C(L)$ function]{%
		Numerical construction of the $C(L)$ function from $S_\mrm{RT}$ by considering $
			\sqrt{
				\frac{1}{2G}
				L' \pdv{S}{L'}
			}
		$. What's happening? I have no idea. \sidenote{Update: it seems that this only holds approximately, perhaps it's simply a numerical coincidence. }
	}
	\end{figure}
\subsubsection{The TsT deformed geodesic}
	If we take the usual geodesic (semi-circle) in the undeformed background, and consider a TsT transformation, we get the TsT deformed geodesic $\gamma_\mrm{TsT}$. This also coincides with the \textit{string frame} geodesic, which minimizes the string frame length without the dilaton factor, instead of the Einstein frame length. See \texttt{notesEETTbar.pdf} for a description of this process.

	However, the on-shell action should still be the Einstein frame length; fortunately, this time the result is nice and clean:
	\begin{equation}
		S_E[\gamma_\mrm{TsT}(z0;\epsilon)]
		= \frac{1}{2}
			\frac{\sqrt{z_0^2 + 2\lambda}}{z_0}
			\pqty{
				\frac{\lambda}{\epsilon^2}
				\frac{\sqrt{z_0^2 - \epsilon^2}}{z_0}
				+ \frac{z_0^2 + \lambda}{z_0^2}
				\mop{arctanh}
					\frac{\sqrt{z_0^2 - \epsilon^2}}{z_0}
			}
	\end{equation}
	Unfortunately, the result is highly divergent: $\sim \frac{1}{\epsilon^2}$, even worse than the $\log \frac{1}{\epsilon}$ divergence of a CFT. 
	\sidenote{It might be possible to "renormalize" out the $\sim \frac{1}{\epsilon^2}$ divergence} which might also help us recover a more reasonable on-shell action.
